---
title: "nonprobsampling: An R Package for Inference from Nonprobability Samples Using Multiple Reference Surveys"
date: "2026-07-11"
abstract: >
  We introduce nonprobsampling, an R package for finite population
  inference from nonprobability samples using multiple probability
  reference surveys. The package is built on the generalized estimating
  equations framework of Landsman et al. (2026), where pseudo-weights for
  the nonprobability sample are constructed by aligning its auxiliary
  information with corresponding design-weighted information from
  probability reference surveys. The primary method implemented is
  multi-reference calibration, which enables the integration of auxiliary
  information from multiple reference surveys when no single reference
  survey contains all variables needed for the participation model.
  One-reference methods, including raking ratio calibration, the adjusted
  logistic propensity (ALP) method, and the Chen--Li--Wu (CLW) method, are
  implemented as special cases of the same framework. We describe the
  methodological foundations and package workflow, evaluate performance
  through simulation under various sampling and covariate-coverage
  scenarios, and illustrate the methods using example surveys demonstrating           multi-reference analyses.
output:
  rjtools::rjournal_article
bibliography: article.bib
header-includes:
  - \fancyfoot{}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.path = "figures/"
)
```

# Introduction

Probability surveys are widely regarded as the gold standard for finite
population inference in survey research [@ElliottValliant2017;
@CornesseEtAl2020]. Their sampling designs explicitly control how
participants are selected, enabling valid generalization from the sample
to the target population. In practice, however, probability surveys are
often costly, time-consuming, and difficult to sustain over time
[@Mindell2015]. As a result, researchers have increasingly turned to
nonprobability samples collected through online platforms, clinical
settings, volunteer panels, or other convenience-based recruitment
channels [@BakerEtAl2013; @HaysEtAl2015]. Although these samples are
easier and faster to collect, their underlying participation mechanisms
are typically unknown. When participation is related to variables that
also affect the outcome of interest, unadjusted estimates of population
means or prevalence may be substantially biased [@Groves2006;
@Bethlehem2010].

A common strategy to correct for participation bias in the
nonprobability sample is to integrate it with an external probability
reference survey that share the same target population. Auxiliary
variables observed in both data sources are then used to estimate
participation probabilities in the nonprobability sample. The inverses
of these estimated probabilities serve as pseudo-weights that function
analogously to survey sampling weights in probability surveys. Several
one-reference pseudo-weighing methods have been proposed, including the
adjusted logistic propensity (ALP) method [@Wang2021], the CLW method
[@Chen2020], and the raking ratio calibration method described in
@Landsman2026.

These methods are useful when a single probability reference survey
contains all auxiliary variables needed for the participation model to
estimate the pseudo-weights, but this requirement can be limiting in
practice. Participation may depend on variables from different domains,
such as demographic, health, occupational, or geographic
characteristics, and a single reference survey may not capture all
relevant variables. When relevant auxiliary variables are missing from
the probability reference survey, the participation model may be
misspecified, leading to biased pseudo-weighted estimates. The
multi-reference calibration method addresses this limitation by
integrating auxiliary information across multiple probability reference
surveys, thereby alleviating the need for all participation-related
variables to be observed within a single reference survey @Landsman2026.

This paper introduces `nonprobsampling`, an R package for finite
population inference from nonprobability samples using multiple
probability reference surveys. The package is structured around the
multi-reference calibration framework of @Landsman2026, which serves as
both the main methodological basis and the core source of software
functionality. Within this framework, the one-reference raking ratio
calibration method, as well as the ALP and CLW methods, are implemented
as special cases for comparison and reproducibility.

The remainder of the paper is organized as follows. We first review the
methodological background for pseudo-weight estimation with multiple
probability reference surveys. We then describe the workflow and
implementation details of `nonprobsampling`. Finally, we illustrate
multi-reference analyses using the package's example datasets and
discuss practical considerations for applied use.

# Methodological Background

This section outlines the methodological foundations underlying the
computations implemented in `nonprobsampling`. The complete statistical
framework, regularity conditions, and variance derivations are described
in @Landsman2026. Here, we focus on the estimating equations framework
that underlies the package, the construction of pseudo-weights from
multiple probability reference surveys, and the special case of the
one-reference method implemented within the package.

## Nonprobability Samples and Pseudo-Weights

Let $\mathrm{FP}$ denote a finite target population of size $N$. For
each unit $i \in \mathrm{FP}$, let $y_i$ be the outcome of interest and
let $\mathbf{x}_i$ be a vector of auxiliary variables for $i=1,...,N$.
The target parameter is the finite population mean: $$\mu
=
\frac{1}{N}\sum_{i\in \mathrm{FP}} y_i .$$ When $y_i$ is binary, $\mu$
corresponds to a finite population prevalence.

Suppose that a nonprobability sample $s_c$ is obtained from
$\mathrm{FP}$. Let $\delta_i^c$ be the indicator that unit $i$ is
included in the nonprobability sample, and define the participation
probability as $$\pi_i^c
=
\Pr(\delta_i^c=1 \mid i\in \mathrm{FP}) .$$ Unlike a probability survey,
a nonprobability sample is not generated under a known sampling design.
Therefore, the participation probabilities $\pi_i^c$ are unknown. If
they were known, a standard Hájek estimator of $\mu$ is given by
$$\hat{\mu}
=
\frac{
\sum_{i\in s_c} y_i / \pi_i^c
}{
\sum_{i\in s_c} 1 / \pi_i^c
}.$$ The inverse participation probability $1/\pi_i^c$ plays the role of
a weight for unit $i$ in the nonprobability sample. Since $\pi_i^c$ is
unknown in practice, `nonprobsampling` estimates this quantity using
auxiliary information from one or multiple probability reference
surveys. The inverse of the estimated participation probabilities are
referred to as pseudo-weights.

The validity of pseudo-weights depends on assumptions about the
participation mechanism. A key assumption is that, conditional on the
auxiliary variables included in the participation model, participation
in the nonprobability sample is independent of the outcome:
$$\Pr(\delta_i^c=1 \mid y_i,\mathbf{x}_i)
=
\Pr(\delta_i^c=1 \mid \mathbf{x}_i).$$ This assumption implies that the
auxiliary variables $\mathbf{x}_i$ should explain both participation in
the nonprobability sample and the variation in the outcome $y_i$. A
positivity condition is also required: $$\pi_i^c > 0
\quad
\text{for all } i\in \mathrm{FP}.$$ Thus, every unit in the finite
population is assumed to have a nonzero probability of participating in
the nonprobability sample. Under these assumptions, pseudo-weights can
be used to reduce participation bias and to estimate FP means from the
nonprobability sample.

## Generalized Estimating Equations Framework

Let $s_{p_1},\ldots,s_{p_K}$ denote $K \geq 1$ probability reference
surveys sampled from the same finite population. For reference survey
$s_{p_k}$, let $\delta_i^{(p_k)}$ be the indicator that unit $i$ is
included in the survey, and let $$d_i^{(p_k)}
=
\frac{1}{\pi_i^{(p_k)}}$$ be the known survey weight, where
$\pi_i^{(p_k)}$ is the survey inclusion probability.

Each reference survey may contain only a subset of the auxiliary
information required for the participation model. Let
$\mathbf{x}_i^{(p_k)}$ denote the auxiliary variables contributed by
reference survey $s_{p_k}$, and let $\mathbf{x}_i$ denote the full
auxiliary vector obtained by stacking the components contributed across
all reference surveys so that the conditional independence assumption
holds. The core estimating equations framework is given by:

\begin{equation}
\left\{
\begin{aligned}
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^c
w(\mathbf{x}_i,\boldsymbol{\beta})
h(\mathbf{x}_i^{(p_1)},\boldsymbol{\beta})
-
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^{(p_1)}
d_i^{(p_1)}
h(\mathbf{x}_i^{(p_1)},\boldsymbol{\beta})
&=
\mathbf{0},
\\
&\vdots
\\
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^c
w(\mathbf{x}_i,\boldsymbol{\beta})
h(\mathbf{x}_i^{(p_K)},\boldsymbol{\beta})
-
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^{(p_K)}
d_i^{(p_K)}
h(\mathbf{x}_i^{(p_K)},\boldsymbol{\beta})
&=
\mathbf{0}.
\end{aligned}
\right.
(\#eq:gee-multi)
\end{equation}

Here, $w(\mathbf{x}_i,\boldsymbol{\beta})$ is a weight function,
$h(\cdot)$ is a vector-valued estimating function, and
$\boldsymbol{\beta}$ is a nuisance parameter vector. In each equation,
the difference between the weighted auxiliary information from the
nonprobability sample and the corresponding design-weighted auxiliary
information from one probability reference survey is set to zero.
Solving this system of equations yields $\hat{\boldsymbol{\beta}}$,
which determines the pseudo-weights for the nonprobability sample
through $$\hat{w}_i
=
w(\mathbf{x}_i,\hat{\boldsymbol{\beta}}),
\qquad i\in s_c.$$ The pseudo-weighted estimator of the finite
population mean is then given by \begin{equation}
\hat{\mu}(\hat{\boldsymbol{\beta}})
=
\frac{
\sum_{i \in s_c}
\hat{w}(\mathbf{x}_i,\hat{\boldsymbol{\beta}})y_i
}{
\sum_{i \in s_c}
\hat{w}(\mathbf{x}_i,\hat{\boldsymbol{\beta}})
}.
(\#eq:pseudo-weighted-mean)
\end{equation}

## Multi-Reference Calibration Method

The main method implemented by `nonprobsampling` is the multi-reference
raking ratio calibration method of @Landsman2026. This method is used
when $K>1$ probability reference surveys are supplied. It is tailored
for situations in which no individual reference survey includes all
auxiliary variables required for the participation model, but the union
of the multiple reference surveys does.

The multi-reference calibration method is defined by substituting the
following choices of the weight function and estimating function into
Equation \@ref(eq:gee-multi): $$w(\mathbf{x}_i,\boldsymbol{\beta})
=
\exp(-\boldsymbol{\beta}^{\top}\mathbf{x}_i),
\qquad
h(\mathbf{x}_i^{(p_k)},\boldsymbol{\beta})
=
\mathbf{x}_i^{(p_k)},
\quad k=1,\ldots,K.$$ With these choices, the estimating equation system
aligns the pseudo-weighted auxiliary totals from the nonprobability
sample with the corresponding design-weighted auxiliary totals from
probability reference surveys. Solving this system gives the parameter
estimate $\hat{\boldsymbol{\beta}}_{\mathrm{cal}}$, which determines the
final multi-reference calibration pseudo-weights.

### Cumulative Precalibration *(Variable Selection and Cumulative Precalibration)*

When multiple probability reference surveys are used, the same auxiliary
variable may appear in more than one reference survey. This overlap
requires special handling before pseudo-weight estimation. First, each
auxiliary variable should contribute only one design-weighted total in
the estimation equations framework. Therefore, overlapping variables
must be properly selected from reference surveys. Second, variables that
are not selected from a particular reference survey may still carry
useful information through their relationships with variables that are
unique to that survey. Cumulative precalibration is used to preserve
this information.

In `nonprobsampling`, two ordering rules are available for selecting
overlapping auxiliary variables. The first orders reference surveys by
sample size, from largest to smallest, so that overlapping variables are
preferentially taken from the larger survey. This is the ordering rule
described in @Landsman2026. The rationale is that larger probability
reference surveys are generally expected to provide more informative
weighted totals and are therefore treated as more reliable reference
sources. Under this rule, the largest survey is processed first and
provides the estimate of the finite population size and the weighted
totals for overlapping auxiliary variables.

The second rule follows the order specified by the user. This option is
useful when sample size alone does not determine the preference of
auxiliary information sources. For example, a smaller survey may be
considered more relevant to the target population, may use more
compatible variable definitions, or may be regarded as more reliable for
a particular application. Under this rule, the user should place the
most important reference survey first, since the estimated population
size and the weighted totals for overlapping variables are taken from
that survey. The next most important survey should be placed second, and
so on.

For example, suppose $K=2$ and the auxiliary variables are $x_1$, $x_2$,
and $x_3$. Suppose that $s_{p_2}$ contains $x_1$ and $x_2$, while
$s_{p_1}$ contains $x_1$ and $x_3$. If $s_{p_2}$ is processed first, the
overlapping variable $x_1$ is assigned to $s_{p_2}$.The assigned vectors
are
then$$ \mathbf{x}_i^{(p_2)}=(1,x_{i1},x_{i2})^\top, \qquad \mathbf{x}_i^{(p_1)}=x_{i3}. $$

The full auxiliary vector in the participation model is formed by
stacking these assigned vectors.

Cumulative precalibration is a sequential procedure for adjusting the
survey weights of later reference surveys before the multi-reference
calibration step. The reference surveys are first ordered according to
the ordering rule discussed above, specified by `sp_order` in the
package. The first survey provides the initial target totals, including
the estimated finite population size and the weighted totals of its
auxiliary variables. Each later survey is then calibrated to the current
targets for overlapping auxiliary variables that are included in the
participation model, as well as to the current estimate of the finite
population size. If there are no overlapping auxiliary variables, only
the sum of weights is calibrated, so that the reference surveys
correspond to the same estimated finite population size. After this
step, the calibrated weights from the current survey are used to
estimate totals for auxiliary variables not yet included in the target
vector. These new totals are appended to the target vector before the
next survey is processed. After all reference surveys have been
processed, the accumulated target vector contains auxiliary totals from
all reference surveys and is used in the subsequent multi-reference
calibration step to estimate pseudo-weights for the nonprobability
sample.

Returning to the two-survey example, $x_1$ is assigned to $s_{p_2}$,
while $x_3$ is assigned to $s_{p_1}$. Although $x_1$ from $s_{p_1}$ is
not included directly in the final calibration system, it may contain
useful information through its relationship to $x_3$ within $s_{p_1}$.
Cumulative precalibration adjusts the weights of $s_{p_1}$ so that the
weighted total of $x_1$ and the total survey weight match the
corresponding totals from $s_{p_2}$. The adjusted weights for $s_{p_1}$
are then used when estimating the design-weighted total of $x_3$.

*For example, suppose age is observed in both reference surveys, a
health variable is observed only in the first survey, and an employment
variable is observed only in the second survey. If age is taken from the
first survey in the final calibration equations, the age variable from
the second survey is not directly included again. By calibrating the
second survey weights to match the age totals from the first survey
before the pseudo-weight estimation step, cumulative precalibration
allows the employment variable to contribute information that remains
aligned with the common age distribution.*

Cumulative precalibration in `nonprobsampling` is based on the
overlapping variables included in `p_formula`, not on all overlapping
variables in the raw data. This design keeps the procedure tied to the
specified participation model and avoids removing observations because
of missing values in variables that are not used for pseudo-weight
estimation.

## One-Reference Special Cases

When $K=1$, Equation \@ref(eq:gee-multi) simplifies to a single
estimating equation:\begin{equation}
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^c
w(\mathbf{x}_i,\boldsymbol{\beta})
h(\mathbf{x}_i,\boldsymbol{\beta})
-
N^{-1}\sum_{i\in \mathrm{FP}}
\delta_i^p
d_i
h(\mathbf{x}_i,\boldsymbol{\beta})
=
\mathbf{0}.
(\#eq:one-ref-gee)
\end{equation} The `nonprobsampling` package implements three choices of
$w(\mathbf{x}_i,\boldsymbol{\beta})$ and
$h(\mathbf{x}_i,\boldsymbol{\beta})$ corresponding to raking raio
calibration method, the adjusted logistic propensity (ALP) method, and
the Chen--Li--Wu (CLW) method.

The one-reference **raking ratio calibration** estimator is obtained by
setting $$w(\mathbf{x}_i,\boldsymbol{\beta})
=
\exp(-\boldsymbol{\beta}^{\top}\mathbf{x}_i),
\qquad
h(\mathbf{x}_i,\boldsymbol{\beta})
=
\mathbf{x}_i .$$ This corresponds to the special case of the
multi-reference calibration method with a single probability reference
survey.

The **ALP** method is defined by $$w(\mathbf{x}_i,\boldsymbol{\beta})
=
\exp(-\boldsymbol{\beta}^{\top}\mathbf{x}_i),
\qquad
h(\mathbf{x}_i,\boldsymbol{\beta})
=
\{1+w(\mathbf{x}_i,\boldsymbol{\beta})\}^{-1}
\mathbf{x}_i .$$

The **CLW** method is defined by $$w(\mathbf{x}_i,\boldsymbol{\beta})
=
1+\exp(-\boldsymbol{\beta}^{\top}\mathbf{x}_i),
\qquad
h(\mathbf{x}_i,\boldsymbol{\beta})
=
\{w(\mathbf{x}_i,\boldsymbol{\beta})\}^{-1}
\mathbf{x}_i .$$

Thus, the one-reference methods are special cases of Equation
\@ref(eq:gee-multi) with $K=1$ and method-specific choices of $w(\cdot)$
and $h(\cdot)$.

## Variance Estimation

Let $z_i$ be a domain indicator, where $z_i=1$ for units in the domain
and $z_i=0$ otherwise. The overall mean is obtained as the special case
where $z_i=1$ for all units in the nonprobability sample. The
pseudo-weighted domain mean estimator is $$\hat{\mu}_z
=
\frac{
\sum_{i\in s_c}
\hat{w}_i z_i y_i
}{
\sum_{i\in s_c}
\hat{w}_i z_i
}.$$

Variance estimation in `nonprobsampling` is based on the Taylor
linearization approach described in @Landsman2026. The variance
estimator accounts for variation from the nonprobability sample,
uncertainty from estimating $\boldsymbol{\beta}$, and sampling variation
in the probability reference survey or surveys.

For the raking ratio calibration estimator with $K$ reference surveys,
the Taylor-linearized variance estimator can be written as
$$\widehat{\mathbb{V}}
\{\hat{\mu}(\hat{\boldsymbol{\beta}}_{\mathrm{cal}})\}
=
T_2^{-2}
\left[
\sum_{i\in s_c}
w(\mathbf{x}_i,\hat{\boldsymbol{\beta}})
\{w(\mathbf{x}_i,\hat{\boldsymbol{\beta}})-1\}
(y_i z_i-\hat{\mu}z_i-\hat{\mathbf{b}}^\top \mathbf{x}_i)^2
+
\hat{\mathbf{b}}^\top \widehat{\mathbf{D}}\hat{\mathbf{b}}
\right],$$ where $$T_2
=
\sum_{i\in s_c}
w(\mathbf{x}_i,\hat{\boldsymbol{\beta}})z_i,$$

$$\hat{\mathbf{b}}^\top
=
\left[
\sum_{i\in s_c}
w(\mathbf{x}_i,\hat{\boldsymbol{\beta}})
z_i
(y_i-\hat{\mu})
\mathbf{x}_i^\top
\right]
\left[
\sum_{i\in s_c}
w(\mathbf{x}_i,\hat{\boldsymbol{\beta}})
\mathbf{x}_i\mathbf{x}_i^\top
\right]^{-1},$$ and $$\widehat{\mathbf{D}}
=
\operatorname{blockdiag}
\left(
\widehat{\mathbf{D}}_1,\ldots,\widehat{\mathbf{D}}_K
\right),
\qquad
\widehat{\mathbf{D}}_k
=
\widehat{\mathbb{V}}_{p_k}
\left(
\sum_{i\in s_{p_k}}
d_i^{(p_k)}
\mathbf{x}_i^{(p_k)}
\right).$$

Here, $\widehat{\mathbf{D}}_k$ denotes the estimated variance-covariance
matrix of the design-weighted auxiliary-variable totals from reference
survey $s_{p_k}$, computed under its probability sampling design. With
multiple reference surveys, these design-based variance-covariance
matrices are combined into the block-diagonal matrix
$\widehat{\mathbf{D}}$. When $K=1$, this construction reduces to the
one-reference case, where $\widehat{\mathbf{D}}=\widehat{\mathbf{D}}_1$.

The first term inside the brackets represents the nonprobability sample
component, evaluated under a Poisson sampling approximation. The second
term, $\hat{\mathbf{b}}^\top \widehat{\mathbf{D}}\hat{\mathbf{b}}$,
captures sampling variation in the auxiliary-variable totals estimated
from the probability reference survey or surveys. In the package, users
provide reference surveys as survey design objects, and the design
information stored in these objects is used to compute
$\widehat{\mathbf{D}}_k$.

# The R package

The `nonprobsampling` package provides tools for finite population
inference from a nonprobability sample using auxiliary information from
probability reference survey(s). The package implements pseudo-weighing
methods based on the generalized estimating equations framework
described in the Method section. Its main purpose is to help users
estimate pseudo-weights for units in a nonprobability sample and then
use those pseudo-weights to estimate population means (prevalence) and
their corresponding standard errors.

The package is designed for settings in which the outcome of interest is
observed in the nonprobability sample, while auxiliary variables related
to participation are available in the nonprobability sample and in one
or multiple probability reference surveys. The main functionality of
`nonprobsampling` is multi-reference calibration, which supports
pseudo-weight estimation when auxiliary variables are distributed across
multiple reference surveys. One-reference methods are implemented as
special cases in this framework, including the raking ratio calibration
method of @Landsman2026, the adjusted logistic propensity (ALP) method
of @Wang2021, and the Chen--Li--Wu (CLW) method of @Chen2020.

## Package design

### Design goals and intended users

The package was developed with applied researchers, survey
statisticians, epidemiologists, social scientists, and data analysts in
mind, particularly those seeking to make inferences from nonprobability
samples. The intended user is expected to be familiar with data frames,
model formulas, and basic survey-weighted analysis in R, but not
necessarily with the internal estimating equations behind the
implemented methods.

The design has three main goals. First, the package provides a direct
workflow from input data to pseudo-weighted estimates. Users do not need
to manually construct estimating equations, compute calibration totals,
or derive variance components. Second, the package integrates existing
survey analysis tools in R from the `survey` package. Reference surveys
are supplied as `survey` design objects rather than as plain data
frames, allowing users to retain survey design information without
manually extracting it. Third, the package is designed for realistic
practical settings in which auxiliary information related to
participation is distributed across multiple reference surveys, while
still supporting the conventional setting in which variables needed for
the participation model are only available from a single reference
survey.

### Integration with the `survey` package

The package relies on the `survey` package to represent probability
reference surveys. This design avoids introducing a new representation
for complex survey designs and allows users to work with familiar
`survey` objects. Reference surveys can be created using
`survey::svydesign()` when the survey design is represented by primary
sampling units, strata, and weights. They can also be created using
`survey::svrepdesign()` when replicate weights are provided.

This integration is important for variance estimation as the uncertainty
of pseudo-weighted mean partially depends on the sampling designs of the
reference surveys. By requiring reference surveys to be supplied as
`survey` design objects, `nonprobsampling` can use the design
information retained in those objects directly. This makes the interface
consistent with standard survey analysis in R and reduces the risk that
users accidentally discard design information when preparing reference
surveys.

### Input objects, formula interface, and workflow

The package workflow is organized around a two-stage design. First,
`est_pw()` estimates pseudo-weights by fitting the participation model
with auxiliary information from the nonprobability sample and multiple
probability reference surveys. Second, `pwmean()` uses the fitted
pseudo-weight object to estimate pseudo-weighted means (prevalence) and
their corresponding standard errors for specified outcomes. This
separation reflects both the statistical structure of the methods and
the intended software workflow: the participation model is fitted once
to construct pseudo-weights, and the resulting object can then be reused
for different outcomes or domains without refitting the model.

The data sources are supplied to `est_pw()` through the `data` argument,
which must be a list. The first element must be the nonprobability
sample, supplied as a data frame. The remaining elements must be
probability reference surveys represented as survey design objects
created with the `survey` package. Thus, an analysis with two reference
surveys has the form

``` r
data = list(sc, ref1_design, ref2_design)
```

where `sc` is the nonprobability sample and `ref1_design` and
`ref2_design` are survey design objects.

The participation model is specified through the argument `p_formula`.
With multiple reference surveys, `p_formula` is a list of one-sided
formulas, with one formula corresponding to each reference survey. For
example,

``` r
p_formula = list(
~ agecat + race + education + pros_enlarged,
~ agecat + race + BMI + diabetes
)
```

specifies that the first reference survey contributes information on
`agecat`, `race`, `education`, and `pros_enlarged`, whereas the second
reference survey contributes information on `agecat`, `race`, `BMI`, and
`diabetes`. The first formula corresponds to the first reference survey
design object after `sc`, the second formula corresponds to the second
reference survey design object, and so on.

A typical multi-reference workflow call has the following form:

``` r
fit_multi <- est_pw(
data = list(sc, ref1_design, ref2_design),
p_formula = list(
~ agecat + race + education + pros_enlarged,
~ agecat + race + BMI + diabetes
),
method   = "multi",
sp_order = "size",
precali  = TRUE
)

est_multi <- pwmean(fit_multi, y = "psa_level")
```

In this example, `education` and `pros_enlarged` are contributed by the
first reference survey, whereas `BMI` and `diabetes` are contributed by
the second reference survey. The variables `agecat` and `race` appear in
both formulas and are therefore treated as overlapping auxiliary
variables. The handling of overlapping variables follows the procedure
discussed in the Cumulative Precalibration section.

The argument `sp_order` controls how multiple reference surveys are
ordered. With `sp_order = "size"`, reference surveys are ordered by
sample size, from largest to smallest. With `sp_order = "given"`, they
are processed in the order supplied by the user. The argument `precali`
controls whether cumulative precalibration is applied before the main
multi-reference calibration step. When `precali = TRUE`, which is the
default, overlapping variables specified in `p_formula` are used to
align weighted totals across reference surveys before the pseudo-weights
estimation step.

The one-reference workflow is a special case of the same interface. When
only one probability reference survey is supplied, the user provides the
nonprobability sample and one survey design object in `data`, and
`p_formula` is a single one-sided formula rather than a list of
formulas.

``` r
fit_one <- est_pw(
data      = list(sc, ref1_design),
p_formula = ~ agecat + race + education + BMI,
method    = "calibration"
)

est_one <- pwmean(fit_one, y = "psa_level")
```

This workflow is useful when a single probability reference survey
contains all auxiliary variables included in the participation model.
Users may choose the raking ratio calibration method, the ALP method, or
the CLW method by changing the `method` argument.

After pseudo-weights have been estimated, `pwmean()` provides a common
outcome-estimation interface. For continuous outcomes, `pwmean()`
estimates a population mean. For binary 0/1 outcomes supplied as
numeric, the same mean estimator corresponds to a prevalence. For
categorical outcomes supplied as a factor, the function returns
prevalence estimates for each observed category. All estimates are
returned with corresponding standard errors and 95% confidence
intervals.

``` r
mean_est <- pwmean(fit_multi, y = "psa_level")

prev_est <- pwmean(fit_multi, y = "BMI")
```

Domain-specific estimates can be obtained by passing a categorical
domain variable supplied as a factor or 0/1 binary to `zcol`. In this
case, `pwmean()` computes the pseudo-weighted estimate separately within
each BMI subgroup.

``` r
domain_est <- pwmean(
fit_multi,
y    = "psa_level",
zcol = "BMI"
)
```

The same fitted pseudo-weights are reused across these calls. Users do
not need to refit the participation model for each outcome or domain.
This is useful in applied analyses where several outcomes are studied
using the same set of auxiliary variables for participation adjustment.
The fitted `"pw_fit"` object stores the method-specific quantities
needed for point and variance estimation, so users do not need to pass
the reference surveys again when calling `pwmean()`.

If `p_formula = NULL`, the package constructs a default participation
model based on variables shared by name between the nonprobability
sample and each reference survey. This default is useful for exploratory
analyses and examples, but in applied work we recommend explicitly
specifying `p_formula`. Explicit formulas make the analysis reproducible
and allow the user to decide which auxiliary variables are plausible
determinants of participation.

If `method = NULL`, `est_pw()` selects the pseudo-weighting method
automatically based on the number of reference surveys supplied. When
multiple reference surveys are provided, it uses `"multi"` for the
multi-reference calibration method. When a single reference survey is
provided, it uses `"calibration"` for the one-reference calibration
method. These defaults correspond to the methods evaluated most
favorably in the simulation study of Landsman et al. (2026) and are
recommended for general use, while users can still explicitly specify
any other supported method when needed.

### Output objects and S3 methods

The function `est_pw()` returns an object of class `"pw_fit"`. This
object contains both user-facing results and internal quantities needed
by `pwmean()`. Important components include the estimated
pseudo-weights, the updated nonprobability sample with a pseudo-weight
column, estimated participation model coefficients, solver diagnostics,
and missing-data information.

The updated nonprobability sample is stored in the component
`sc_updated`. By default, the pseudo-weight column added to this data
frame is named `pseudo_wts`, and users can change this name through the
`sc_wname` argument. The estimated pseudo-weights are also stored
separately as a numeric vector in the component `pseudo_weights`. Solver
information is stored in `solver_diagnostics`, including the solver
name, termination code, solver message, number of iterations, and the
maximum absolute estimating equations value at convergence.

The function `pwmean()` returns an object of class `"pwmean"` for
numeric and binary outcomes. For categorical outcomes, it returns an
object of class `"pwmean_factor"`, which returns category-specific
prevalence estimates. The main component is `estimates`, a data frame
containing the unweighted and pseudo-weighted estimates, their standard
errors, and 95% confidence intervals. For numeric outcomes, the estimate
is a mean. For binary 0/1 outcomes, the estimate corresponds to
prevalence. For categorical outcomes, each category is internally
converted into an indicator, and the function returns prevalence
estimate for each observed category.

The package provides S3 methods for printing and summarizing fitted
objects. The `print()` method for `"pw_fit"` objects gives a compact
overview of the fitted pseudo-weight object, including the method,
participation model size, convergence status, and pseudo-weight summary.
The `summary()` method provides more detailed output, including
coefficients and solver diagnostics. The `print()` and `summary()`
methods for `"pwmean"` and `"pwmean_factor"` objects display the
unweighted and pseudo-weighted estimates, standard errors, and
confidence intervals. The package also provides `na.action()` methods to
retrieve missing data information of objects returned by `est_pw()` and
`pwmean()`.

## Practical considerations

### Variable harmonization

Variable harmonization is required before using `est_pw()`. The package
matches variables by name, not by meaning. Therefore, a variable called
`agecat` in the nonprobability sample and `age_group` in a reference
survey will not be treated as the same variable unless the user renames
one of them. Similarly, factor levels must be compatible across data
sources. Users should harmonize shared variables so that they have the
same names, compatible categories, and the same definitions.

Harmonization is especially important in multi-reference analyses. Some
variables may appear in multiple reference surveys, while others appear
in only one. The user should decide which variables are comparable and
whether differences in survey timing, question wording, or target
population definitions are acceptable. The package checks variable
availability and factor compatibility, but it cannot determine whether
two variables have the same definition.

Continuous variables should also be harmonized. For example, income
measured in dollars in one survey and thousands of dollars in another
survey should be rescaled before estimation. Similarly, biomarkers,
durations, or expenditure variables should use comparable units. If the
scale of a continuous auxiliary variable differs across data sources,
the fitted participation model and resulting pseudo-weights may be
misleading.

### Missing data handling

The package handles missing data in two stages. The first stage occurs
in `est_pw()` and applies to variables used in the participation model.
The second stage occurs in `pwmean()` and applies to the outcome
variable and, if supplied, the domain variable.

By default, `est_pw()` uses `na.action = stats::na.omit`. Rows with
missing values in the participation model variables are removed before
pseudo-weights are estimated, so these rows are not included in either
the `sc_updated` data frame or the `pseudo_weights` vector of the
returned `"pw_fit"` object. Information about the omitted rows is
recorded in the fitted object and can be inspected through the summary
output or by calling `na.action()` on the fitted object.

Alternatively, users may use `na.action = stats::na.exclude`. In this
case, observations in the nonprobability sample that are excluded from
pseudo-weight estimation remain in the returned `sc_updated` data frame,
but their pseudo-weight values are recorded as `NA`. This option is
useful when users want to preserve row alignment with the original
nonprobability sample for later merging, checking, or imputation.

After pseudo-weights have been estimated, `pwmean()` applies
missing-data handling separately to the outcome variable and, if
supplied, the domain variable specified by `zcol`. By default,
`pwmean()` uses `na.action = stats::na.omit`, so observations with
missing outcome or domain values are excluded from that specific outcome
analysis. This separate handling reflects the two-stage workflow:
missing participation model variables affect only pseudo-weight
estimation, while missing outcome or domain values affect only the
outcome-specific estimation.

### Solver control and diagnostics

Pseudo-weight estimation requires solving estimating equations. The
function `pw_solver_control()` creates a control object for this
numerical step. The default solver is `"nleqslv"`, and users can modify
convergence tolerances, the maximum number of iterations, tracing
behavior, and solver-specific options.

For most applications, the default settings are sufficient. However,
convergence can be more difficult to reach when the participation model
contains many parameters, sparse factor levels, strong collinearity, or
auxiliary variables with limited overlap across data sources. In such
cases, users may adjust the solver controls:

``` r
ctrl <- pw_solver_control(
method = "Broyden",
xtol   = 1e-10,
ftol   = 1e-6,
maxit  = 100,
trace  = TRUE
)
```

The fitted object stores solver diagnostics. Users should inspect the
termination message, the number of iterations, and the final maximum
absolute value of the estimating equations. Nonconvergence may indicate
that the participation model is too complex for the available data, that
factor levels are sparse, or that the nonprobability sample and
reference survey have poor overlap in the auxiliary-variable space. In
such situations, users should consider simplifying the participation
model, combining sparse categories, checking variable harmonization, or
evaluating whether the reference survey is appropriate for the target
population.

### Replicate-weight reference surveys

Some probability reference surveys provide replicate weights rather than
primary sampling unit and stratum identifiers. The package supports this
workflow through `survey::svrepdesign()`. A replicate-weight survey
design object can be supplied to `est_pw()` in the same way as a
standard survey design object.

This design allows the package to use replicate-weight information
supplied by survey data providers. In applied work, users should use the
replicate weights and replicate-weight type described in the
documentation for the reference survey. A worked example using the
package data is provided in Illustration section

## Implementation details

### Dependencies

The package imports a small set of dependencies, keeping installation
lightweight and making the package easy to use in applied environments.
The `survey` package is used to represent and process probability
reference surveys. This allows users who already analyze probability
surveys with `survey` to incorporate `nonprobsampling` into their
existing workflow. The `nleqslv` package is used to solve the nonlinear
estimating equations for pseudo-weight estimation. Functions from
`stats` are used for formula handling, model matrix construction,
missing-data handling, and related base statistical operations. The
package also uses `utils` for standard utility functions.

### Variance estimation

The package reports standard errors and confidence intervals for
pseudo-weighted estimates. Analytic variance estimation is based on
Taylor linearization and accounts for uncertainty from pseudo-weight
estimation as well as the sampling designs of the probability reference
surveys. For reference surveys represented by survey design objects, the
design information is taken from the survey design object. The variance
estimator depends on the method-specific quantities saved during
pseudo-weight estimation. This is why `pwmean()` takes the fitted object
as its argument rather than taking pseudo-weights alone.

### Computational considerations

The computational cost of `est_pw()` is determined primarily by the
numbers of observations in the nonprobability sample and reference
surveys, the number of reference surveys, the dimensions of the expanded
participation model matrices, and the number of iterations required by
the nonlinear equation solver. Categorical variables with many levels,
interaction terms, and polynomial terms can substantially increase the
model-matrix columns the participation model parameters. Because
est_pw() repeatedly evaluates the estimating equations and their
Jacobian matrices, increases in model-matrix columns can substantially
increase both computation time and memory use. Large or poorly
conditioned model matrices may also result in rank deficiency, numerical
instability, or slower convergence.

The reference survey designs contribute additional computational cost
because `est_pw()`calculates design-based covariance matrices for the
quantities required in subsequent variance estimation. This cost depends
on the sample sizes, the numbers of model-matrix columns, and the survey
design representations; replicate-weight designs may be particularly
demanding when they contain many replicate weights. In multi-reference
analyses, model matrices, survey-weighted totals, and design-based
covariance blocks are constructed for every reference survey. When
`precali = TRUE`, which is the default, the reference survey weights are
also recalibrated sequentially before pseudo-weight estimation. The cost
of this cumulative precalibration increases with the number and size of
the reference surveys and the number of shared auxiliary variables.

The current implementation uses in-memory R objects and dense model
matrices and is therefore intended primarily for typical applied
survey-analysis settings. Very large samples or high-dimensional
participation models may require substantial memory and computation. For
such analyses, users should inspect the expanded model-matrix
dimensions, retain substantively important variables, combine sparse
factor levels, and avoid unnecessary interaction or polynomial terms
before fitting. Solver diagnostics should be examined after fitting to
assess convergence.

## Related R packages

Several R packages support inference from nonprobability samples. The
closest packages to `nonprobsampling` are `NonProbEst` and `nonprobsvy`.
All three packages address finite population inference from
nonprobability samples by using auxiliary information from external
sources to reduce participation bias. However, they differ in
methodological scope, interface design, and support for analyses using
multiple reference surveys. Table \@ref(tab:related-packages) summarizes
the main differences between the three packages.

```{r related-packages, echo=FALSE}
related_packages <- data.frame(
  Feature = c(
    "Number of reference surveys",
    "Cumulative precalibration",
    "Outcome types",
    "Estimation scope",
    "Survey design integration",
    "Variance estimation method",
    "Core methods",
    "User workflow",
    "Main focus"
  ),
  nonprobsampling = c(
    "Multiple or one",
    "Supported for aligning reference surveys",
    "Continuous, binary, and categorical",
    "Overall and domain-specific",
    "PSU-based and replicate-weight survey designs",
    "Analytical",
    "Multi-reference calibration; raking-ratio calibration; ALP; CLW",
    "est_pw() followed by pwmean()",
    "Pseudo-weighted estimation using multiple reference surveys"
  ),
  nonprobsvy = c(
    "One",
    "Not available",
    "Continuous and binary",
    "Overall",
    "PSU-based survey designs",
    "Analytical and resampling-based",
    "IPW; calibrated IPW; mass imputation; doubly robust estimation",
    "nonprob()",
    "General nonprobability sample inference"
  ),
  NonProbEst = c(
    "One",
    "Not available",
    "Continuous and binary",
    "Overall",
    "No support",
    "Resampling-based",
    "Calibration; PSA; matching; model-based, model-assisted, and model-calibrated estimation",
    "Method-specific functions",
    "Several estimators for nonprobability samples"
  ),
  check.names = FALSE
)

related_packages_output <- knitr::kable(
  related_packages,
  format = if (knitr::is_latex_output()) "latex" else "html",
  caption = paste(
    "Comparison of nonprobsampling with related R packages for",
    "nonprobability sample inference."
  ),
  booktabs = TRUE,
  align = c("l", "l", "l", "l")
)

if (knitr::is_latex_output()) {
  related_packages_output <- kableExtra::column_spec(
    related_packages_output,
    column = 1,
    width = "3.3cm"
  )
  related_packages_output <- kableExtra::column_spec(
    related_packages_output,
    column = 2:4,
    width = "3.9cm"
  )
  related_packages_output <- kableExtra::row_spec(
    related_packages_output,
    row = 1:9,
    extra_latex_after = "\\addlinespace[2pt]"
  )
  kableExtra::kable_styling(
    related_packages_output,
    latex_options = "hold_position",
    font_size = 10
  )
} else {
  kableExtra::kable_styling(
    related_packages_output,
    bootstrap_options = c("striped", "condensed", "responsive"),
    full_width = TRUE
  )
}
```

The `NonProbEst` package provides various methods for estimation in
nonprobability surveys, including calibration to known population
totals, propensity score adjustment with a probability reference survey,
statistical matching, model-based estimation, model-assisted estimation,
and model-calibrated estimation. It also implements machine-learning
methods for propensity score estimation and outcome prediction. The
`nonprobsvy` package also provides several methods with a unified
interface. The package supports calibrated inverse probability
weighting, mass imputation, matching-based prediction, nonparametric
prediction, and doubly robust estimation. It also provides analytical
variance estimation, variable selection, and balance checking.

However, both `NonProbEst` and `nonprobsvy` use reference information
from a single source at a time. In practice, however, accurate
pseudo-weight estimation depends on including auxiliary variables that
are related to participation in the nonprobability sample, and no single
probability reference survey will necessarily contain all such
variables. Important demographic, socioeconomic, health, behavioral, or
occupational variables are often distributed across different reference
surveys.

This is the purpose for which `nonprobsampling` is developed. Its main
contribution is the multi-reference calibration method, which allows
users to combine auxiliary information from multiple probability
reference surveys. By allowing different reference surveys to contribute
different parts of the participation model, the package supports
analyses that would be inaccurate with methods limited to one reference
source. When only one reference survey is available, `nonprobsampling`
supports one-reference pseudo-weighing methods, including raking ratio
calibration, ALP, and CLW, as special cases of the same framework. The
package therefore provides a consistent workflow for both the
conventional one-reference setting and a more practically realistic
multi-reference setting.

The packages also differ in their support for categorical prevalence
estimation and subpopulation estimation. All three packages can handle
binary outcomes. `NonProbEst` supports proportion estimation for binary
indicators, and `nonprobsvy` supports binary outcomes through binomial
outcome models, but neither package provides multi-level categorical
prevalence estimation. By contrast, in `nonprobsampling`, `pwmean()` can
accept a categorical outcome as a factor and return category-specific
prevalence estimates. Besides, neither `NonProbEst` nor `nonprobsvy`
provides support for subgroup estimation in their workflow.
`nonprobsampling` provides this functionality through the `zcol`
argument in `pwmean()`, allowing users to estimate pseudo-weighted means
(prevalence) and their corresponding standard errors within subgroups
rather than only for the full sample. This is useful in practice because
applied analyses often require computation and comparisons across
population domains, such as age groups, sex, race, or health-risk
groups. A method that only returns an overall estimate may hide
important subgroup differences, whereas domain-specific estimates allow
users to report results at the level most relevant to applied research
questions.

The packages also differ in their intended software workflow. In
`NonProbEst`, users often move through a sequence of separate functions
for constructing weights, estimating target quantities, and computing
intervals. In `nonprobsvy`, the main workflow is centered on a single
model-fitting function, `nonprob()`, whose behavior is controlled by
specifying participation, outcome, and target formulas. In
`nonprobsampling`, the workflow is deliberately split into two steps,
separating pseudo-weights estimation from outcome estimation, so the
same fitted pseudo-weights can be reused for multiple outcomes without
refitting the participation model.

The three packages also differ in how they use survey design
information. `NonProbEst` is not integrated with `survey`; users supply
probability reference surveys as ordinary data frames. `nonprobsvy` and
`nonprobsampling` both integrate with `survey`, but they differ in the
types of survey design objects supported for probability reference
surveys. `nonprobsvy` supports reference surveys from standard
multistage sampling designs, supplied as `survey.design` objects. In
contrast, `nonprobsampling` accepts probability reference surveys as
either a `survey.design` object or a `svyrep.design` object. This
feature is useful when reference surveys provide replicate weights
rather than full design information. Such replicate-weight formats are
commonly used by statistical agencies such as Statistics Canada, the
U.S. Census Bureau, and the Australian Bureau of Statistics.

`nonprobsampling` should be viewed as complementary to `NonProbEst` and
`nonprobsvy`. Users who need different types of nonprobability-sample
estimators may find `nonprobsvy` or `NonProbEst` more appropriate. Users
whose main goal is to construct pseudo-weights for a nonprobability
sample, especially when the relevant auxiliary variables are split
across different reference surveys, are the primary audience for
`nonprobsampling`. The package's contribution is therefore not breadth
across all nonprobability sample methods, but a focused and reproducible
implementation of multi-reference pseudo-weighted estimation, with
direct support for standard and replicate-weight survey design objects,
multi-level categorical prevalence estimation, and domain-specific
summaries.

# Simulation

This simulation study evaluates the performance of the multi-reference
calibration method for estimating prevalence in a multi-level
categorical outcome setting. The data generation and sampling design
follows Section 6 of Landsman et al. (2026). The original study used
continuous and binary outcomes, the present study extends the simulation
framework to a three-level categorical outcome with three population
domains. Performance is evaluated for the prevalence of each outcome
category within each domain. The study also compares the multi-reference
calibration estimator with and without cumulative precalibration.

## Finite population and target parameters

A finite population of size $N=500{,}000$ is generated and held fixed
across simulation replications. The finite population contains four
auxiliary variables with one categorical domain variable. The domain
variable $x_1$ has three levels, denoted by A, B, and C, and is
generated according to

$$\Pr(x_{1i}=\mathrm{A})=0.50,\qquad
\Pr(x_{1i}=\mathrm{B})=0.25,\qquad
\Pr(x_{1i}=\mathrm{C})=0.25.$$

Let $I_{Bi}=\mathbb{I}(x_{1i}=\mathrm{B})$ and
$I_{Ci}=\mathbb{I}(x_{1i}=\mathrm{C})$. Three continuous auxiliary
variables are generated as

$$\begin{aligned}
x_{2i} &= \nu_{2i}+0.3I_{Bi}+0.3I_{Ci},\\
x_{3i} &= \nu_{3i}+0.2(I_{Bi}+I_{Ci}+x_{2i}),\\
x_{4i} &= \nu_{4i}+0.1(I_{Bi}+I_{Ci}+x_{2i}+x_{3i}),
\end{aligned}$$

where $\nu_{2i}\sim\mathrm{Unif}(0,2)$, $\nu_{3i}\sim\mathrm{Exp}(1)$,
and $\nu_{4i}\sim\chi^2(4)$.

A latent continuous outcome variable is generated according to

$$y_i^*
=
-0.5I_{Bi}-I_{Ci}-x_{2i}+x_{3i}+x_{4i}
+\epsilon_i,$$

where $\epsilon_i\sim N(0,1)$. The observed outcome $y_i$ is a
three-level categorical variable with levels Low, Medium, and High. The
category labels are assigned based on the empirical 35th and 70th
percentiles of $y^*$ in the finite population:

$$y_i =
\begin{cases}
\mathrm{Low}, & y_i^*\leq c_1,\\
\mathrm{Medium}, & c_1<y_i^*\leq c_2,\\
\mathrm{High}, & y_i^*>c_2,
\end{cases}$$

where $c_1$ and $c_2$ denote the 35th and 70th percentiles of $y^*$.
Therefore, the marginal finite population prevalence of Low, Medium, and
High are approximately 0.35, 0.35, and 0.30. Since the domain variable
$x_1$ affects the outcome distribution, the category prevalence differ
across the three domains.

The parameters of interest are the domain-specific prevalence for each
outcome category. For domain $d$ and outcome category $\ell$, the target
parameter is defined as

$$P_{d\ell}
=
\frac{
\sum_{i\in FP}
\mathbb{I}(x_{1i}=d)
\mathbb{I}(y_i=\ell)
}{
\sum_{i\in FP}
\mathbb{I}(x_{1i}=d)
},$$

where $d\in\{\mathrm{A},\mathrm{B},\mathrm{C}\}$ and
$\ell\in\{\mathrm{Low},\mathrm{Medium},\mathrm{High}\}$. The nine finite
population domain-specific prevalence are calculated before sampling and
treated as the true values for evaluating estimator performance.

## Sampling design

Two probability reference surveys and one nonprobability sample are
independently selected from the finite population in each replication.

The first probability reference survey, $s_{p1}$, is selected by Poisson
sampling with expected sample size $n_{p1}=12{,}500$. Its inclusion
probabilities are

$$\pi_i^{(p1)}
=
\frac{n_{p1}q_i}{\sum_{j\in\mathrm{FP}}q_j},
\qquad
q_i=c_1+x_{3i}+0.03y_i^*,$$

where $c_1$ is chosen so that the ratio of the largest to the smallest
value of $q_i$ is 20.

The second probability reference survey, $s_{p2}$, is selected by
randomized systematic probability-proportional-to-size sampling with
fixed sample size $n_{p2}=25{,}000$. Its first-order inclusion
probabilities are proportional to

$$z_i=0.05+x_{2i}.$$

The nonprobability sample, $s_c$, is generated independently by Poisson
sampling with expected size $n_c=2{,}500$. The participation
probabilities follow

$$\pi_i^c
=
\exp\left(
\beta_0
+0.18I_{Bi}
+0.18I_{Ci}
+0.18x_{2i}
-0.27x_{3i}
-0.1x_{4i}
\right),$$

where $\beta_0$ is selected so that
$\sum_{i\in\mathrm{FP}}\pi_i^c=2{,}500$.

The outcome $y$ and all four auxiliary variables are observed in the
nonprobability sample. In contrast, $s_{p1}$ contains only $x_1$ and
$x_2$, whereas $s_{p2}$ contains only $x_3$ and $x_4$. The two
probability reference surveys therefore supply complementary auxiliary
information, and neither survey alone contains the complete set of
variables in the participation model.

## Simulation scenarios and estimation

Each simulated dataset was analyzed using the multi-reference
calibration method implemented by `est_pw()`. The participation formula
was

``` r
p_formula = list(
  ~ x1 + x2,
  ~ x3 + x4
)
```

corresponding to the auxiliary variables observed in $s_{p1}$ and
$s_{p2}$. Two precalibration cases were compared:

1.  **No precalibration:** multi-reference calibration was fitted with
    `precali = FALSE`.
2.  **Precalibration:** multi-reference calibration was fitted with
    `precali = TRUE`.

Since there were no auxiliary variables shared between the probability
reference surveys, cumulative precalibration in the second case aligned
only the sum of the survey weights (the estimated finite population
size) before the pseudo-weight estimation step. For `sp_order`, the
package used its default size-based ordering. Therefore, the larger
survey, $s_{p2}$, was processed first, and its estimated population size
was used as the target when precalibrating $s_{p1}$.

For each fitted pseudo-weight object, `pwmean()` was called with
`y = "y"` and `zcol = "x1"`. Since $y$ was a factor, the function
estimated the prevalence of each outcome category. Since $x_1$ was
supplied as the domain variable, separate prevalence estimates were
returned for domains A, B, and C. Each fitted model therefore produced
nine pseudo-weighted prevalence, corresponding standard errors and 95%
confidence interval.

## Evaluation metrics

For each domain $d$, outcome category $\ell$, and precalibration case
$m$, the simulation was repeated $B = 4000$ times. In replication $b$,
let $\widehat{P}_{d\ell m}^{(b)}$ denote the estimated prevalence and
let $\widehat{\mathrm{SE}}_{d\ell m}^{(b)}$ denote its analytic standard
error. The true finite population prevalence is denoted by $P_{d\ell}$.

To evaluate the point estimators, relative bias (RB) and mean squared
error (MSE) were calculated as

$$\mathrm{RB}(\%)
=
\frac{100}{B}
\sum_{b=1}^{B}
\frac{\widehat{P}_{d\ell m}^{(b)}-P_{d\ell}}{P_{d\ell}}.$$

and

$$\mathrm{MSE}\times100
=
\frac{100}{B}
\sum_{b=1}^{B}
\left(\widehat{P}_{d\ell m}^{(b)}-P_{d\ell}\right)^2.$$

To evaluate the analytic variance estimators, the average estimated
variance, multiplied by 100, was calculated as

$${v}\times100
=
\frac{100}{B}
\sum_{b=1}^{B}
\left\{\widehat{\mathrm{SE}}_{d\ell m}^{(b)}\right\}^2$$

The standard error ratio (SER) compares the average estimated standard
error with the empirical standard deviation of the prevalence estimates:

$$\mathrm{SER}
=
\frac{
B^{-1}
\sum_{b=1}^{B}
\widehat{\mathrm{SE}}_{d\ell m}^{(b)}
}{
\mathrm{SD}\left(
\widehat{P}_{d\ell m}^{(1)},\ldots,
\widehat{P}_{d\ell m}^{(B)}
\right)
}.$$

Values of SER near 1 indicate agreement between the analytic and
empirical standard errors.

Finally, coverage probability (CP) was calculated as the proportion of
nominal 95% Wald confidence intervals that contained the true
prevalence:

$$\mathrm{CP}
=
\frac{1}{B}
\sum_{b=1}^{B}
\mathbb{I}\left[
\widehat{P}_{d\ell m}^{(b)}-1.96\widehat{\mathrm{SE}}_{d\ell m}^{(b)}
\leq P_{d\ell}\leq
\widehat{P}_{d\ell m}^{(b)}+1.96\widehat{\mathrm{SE}}_{d\ell m}^{(b)}
\right].$$

CP was reported as a proportion between 0 and 1. The metrics are
reported as `RB(%)`, `v(*100)`, `MSE(*100)`, `SER`, and `CP` in the
simulation results table below.

## Simulation results

Table \@ref(tab:simulation-results) summarizes the performance of the
multi-reference calibration estimators for estimating domain-specific,
category-specific prevalence, with and without cumulative
precalibration.

```{r simulation-results, echo=FALSE}
results_table <- data.frame(
  domain = rep(c("A", "B", "C"), times = 6),
  outcome = rep(
    rep(c("Low", "Medium", "High"), each = 3),
    times = 2
  ),
  precali_case = rep(
    c("No precalibration", "Precalibration"),
    each = 9
  ),
  `RB (%)` = sprintf("%.2f", c(
    -0.2383, -0.2595, -0.2233, 0.1807, 0.1635, 0.2378,
    0.0200, 0.1230, 0.0818, -0.2484, -0.2681, -0.2308,
    0.1840, 0.1648, 0.2414, 0.0256, 0.1320, 0.0902
  )),
  `v (*100)` = sprintf("%.2f", c(
    0.0154, 0.0331, 0.0376, 0.0210, 0.0376, 0.0373,
    0.0201, 0.0418, 0.0397, 0.0154, 0.0331, 0.0376,
    0.0210, 0.0376, 0.0373, 0.0201, 0.0418, 0.0397
  )),
  `MSE (*100)` = sprintf("%.2f", c(
    0.0157, 0.0327, 0.0380, 0.0214, 0.0371, 0.0375,
    0.0206, 0.0417, 0.0405, 0.0156, 0.0324, 0.0378,
    0.0214, 0.0371, 0.0375, 0.0205, 0.0416, 0.0405
  )),
  SER = sprintf("%.2f", c(
    0.9915, 1.0065, 0.9946, 0.9918, 1.0060, 0.9977,
    0.9856, 1.0002, 0.9897, 0.9938, 1.0108, 0.9976,
    0.9914, 1.0065, 0.9982, 0.9887, 1.0014, 0.9894
  )),
  CP = sprintf("%.2f", c(
    0.9462, 0.9492, 0.9505, 0.9515, 0.9472, 0.9512,
    0.9470, 0.9480, 0.9448, 0.9435, 0.9528, 0.9528,
    0.9508, 0.9490, 0.9508, 0.9478, 0.9480, 0.9450
  )),
  check.names = FALSE
)

names(results_table)[1:3] <- c("Domain", "Outcome", "Precalibration")

results_table_output <- knitr::kable(
  results_table,
  format = if (knitr::is_latex_output()) "latex" else "html",
  align = rep("c", 8),
  caption = "Simulation results.",
  booktabs = TRUE,
  row.names = FALSE,
  linesep = ""
)

if (knitr::is_latex_output()) {
  simulation_widths <- c("1cm", "1.2cm", "2.4cm", "1cm", "1cm", "1.25cm", "0.95cm", "0.95cm")
  for (column_index in seq_along(simulation_widths)) {
    results_table_output <- kableExtra::column_spec(
      results_table_output,
      column = column_index,
      width = simulation_widths[[column_index]]
    )
  }
  results_table_output <- kableExtra::row_spec(
    results_table_output,
    row = c(3, 6, 9, 12, 15),
    extra_latex_after = "\\addlinespace[3pt]"
  )
  kableExtra::kable_styling(
    results_table_output,
    latex_options = "hold_position",
    font_size = 8
  )
} else {
  kableExtra::kable_styling(
    results_table_output,
    bootstrap_options = c("striped", "condensed", "responsive"),
    full_width = TRUE
  )
}
```

*Note:* RB is relative bias in percent; $v\times100$ is the average
estimated analytic variance multiplied by 100; MSE$\times100$ is the
mean squared error multiplied by 100; SER is the ratio of the average
estimated standard error to the empirical standard deviation of the
point estimates; and CP is the empirical coverage proportion of nominal
95% Wald confidence intervals.

# Illustration

This section illustrates how `nonprobsampling` can be used with multiple
probability reference surveys using the example datasets included in the
package. Because the basic function workflow has already been
introduced, the emphasis here is on the applied use of the
multi-reference method. Specifically, the section explains why multiple
reference surveys are useful in this example, how the multi-reference
interface is specified in practice, how to interpret the output from
`est_pw()` and `pwmean()`, and how to assess whether the numerical
solver has converged successfully (**satisfactorily**). The section also
evaluates the resulting pseudo-weighted estimates by comparing them with
the unweighted estimates from the nonprobability sample and with the
true finite population means used as benchmarks.

## Example data

```{r load-data, include=FALSE}
library(nonprobsampling)

data("sc")
data("sp1")
data("sp2")
```

The illustration uses the example datasets included in
`nonprobsampling`. The three datasets are summarized in Table
\@ref(tab:data-summary). These datasets are used to demonstrate
estimation of the mean serum prostate-specific antigen (PSA) level among
U.S. men aged 55 years and older. The outcome of interest is `psa_level`
(ng/mL), a continuous measure of serum PSA level stored as a numeric
variable.

### Construction of the `sc` sample

Dr. Landsman: PLEASE ADD HERE

JIAKUN:**,** \*\* WHAT IS CORRECT, DATASET or DATA SET?\*\* (**DATASET
will be used through the paper**)

The dataset `sc` represents the nonprobability sample to be adjusted. It
contains the outcome variable `psa_level`, along with auxiliary
variables related to participation.

<!-- DRAFT REMOVED: ### Probability reference surveys -->

**NEED TO DEFINE NHAMES AND NHIS - GOOD TO INCLUDE WHAT THESE SURVEYS
MEASURE IN GENERAL**

<!-- DRAFT REMOVED: The datasets `sp1` and `sp2` represent probability reference surveys and provide external auxiliary information for estimating pseudo-weights. Notably, psa_level is also available in sp1. Although outcomes are typically observed only in the nonprobability sample, including the outcome in the participation model may help reduce bias when the participation probability also depends on the outcome. -->

<!-- DRAFT REMOVED: For illustration, all shared variables have already been harmonized across the example datasets. Variable names, factor levels, and category definitions are consistent. In user-supplied data, this harmonization must be completed before calling est_pw(), because the package assumes that variables with the same name have compatible meanings and coding across data sources. -->

### Probability reference surveys

The datasets `sp1` and `sp2` represent probability reference surveys and
provide external auxiliary information for estimating pseudo-weights.
The dataset `sp1` is derived from the National Health and Nutrition
Examination Survey (NHANES), a nationally representative survey designed
to collect information on health conditions, health-related behaviors,
and clinical measurements in the U.S. population. The dataset `sp2` is
derived from the National Health Interview Survey (NHIS), a nationally
representative survey that collects information on health status, health
care access, and related characteristics through household interviews.
Both surveys use complex probability sampling designs and provide survey
weights that support design-based population inference.

<!-- DRAFT REMOVED: ### Motivation for the multi-reference method -->

**BRING THIS SECTION BEFORE OR RIGHT AFTER PROBABILITY REFERENCE
SURVEYS? (now it's right after)**

These datasets also illustrate why multiple reference surveys are needed
in practice. Both reference surveys contain shared variables such as
`agecat`, `race`, and `comorbidity`. In addition, the first reference
survey, `sp1`, includes variables such as `education` and
`pros_enlarged`, whereas the second reference survey, `sp2`, includes
variables such as `BMI` and `diabetes`. These additional variables are
likely related to participation. Relying on a single reference survey
may limit the set of available auxiliary variables and lead to less
effective adjustment if important variables are missing. The
multi-reference method combines variables from both surveys when
estimating pseudo-weights for the nonprobability sample, thereby
improving the use of available information.

Notably, `psa_level` is also available in `sp1`. Although outcomes are
typically observed only in the nonprobability sample, including the
outcome in the participation model may help reduce bias when
participation probabilities depend on the outcome or outcome-related
characteristics.

For illustration, all shared variables have already been harmonized
across the example datasets. Variable names, factor levels, and category
definitions are consistent. In user-supplied data, this harmonization
must be completed before calling `est_pw()`, because the package assumes
that variables with the same name have compatible meanings and coding
across data sources.

```{r data-summary, echo=FALSE}
dataset_info <- data.frame(
  Dataset = c("sc", "sp1", "sp2"),
  Type = c(
    "Nonprobability sample",
    "First probability reference survey",
    "Second probability reference survey"
  ),
  Source = c(
    "Synthetic sample generated from an NHANES-based finite population",
    "NHANES 1999--2010",
    "NHIS 1997--2008"
  ),
  Sample_size = c(nrow(sc),nrow(sp1),nrow(sp2)),
  check.names = FALSE
)

knitr::kable(
  dataset_info,
  caption = "Datasets included in the package.",
  align = c("l", "l", "l", "r")
)

```

### Survey design objects

Before pseudo-weight estimation, each probability reference survey needs
to be represented as a `survey` design object. These objects store the
survey design information used by `nonprobsampling` when estimating
pseudo-weights and computing standard errors.

```{r survey-designs}
ref1_design <- survey::svydesign(
  ids     = ~psu_sp1,
  strata  = ~strata_sp1,
  weights = ~wts_sp1,
  data    = sp1,
  nest    = TRUE
)

ref2_design <- survey::svydesign(
  ids     = ~psu_sp2,
  strata  = ~strata_sp2,
  weights = ~wts_sp2,
  data    = sp2,
  nest    = TRUE
)
```

The resulting objects, `ref1_design` and `ref2_design`, are then
supplied to `est_pw()` together with the nonprobability sample, `sc`.

## Multi-reference calibration

The example below is a typical use of the main function `est_pw()`, the
two formulas in `p_formula` correspond to the two reference survey
design objects supplied after `sc`. The formulas specify which auxiliary
variables each reference survey contributes to the participation model.
In general, users should include the relevant variables that are shared
between the nonprobability sample and each reference survey, unless
there is a practical reason to exclude them.

Cumulative precalibration is recommended for most multi-reference
analyses. In this example, with `precali = TRUE`, the package uses the
overlapping variables specified in `p_formula`, such as `agecat`,
`race`, and `comorbidity`, to align the reference surveys before
estimating the final pseudo-weights.

The fitted object returned by `est_pw()` stores the estimated
pseudo-weights and the information needed for subsequent outcome
estimation. The `summary()` output provides detailed information about
how the multi-reference calibration method was implemented, while the
`print()` output gives a shorter overview of the fitted object.

```{r multi-ref-fit}
fit_multi <- est_pw(
  data = list(sc, ref1_design, ref2_design),
  p_formula = list(
    ~ agecat + race + education + pros_enlarged + comorbidity + psa_level + BMI,
    ~ agecat + race + BMI + diabetes + comorbidity
  ),
  method   = "multi",
  sp_order = "size",
  precali  = TRUE
)

summary(fit_multi)
```

In this output, the method is reported as multi-reference calibration,
confirming that pseudo-weights were estimated from two probability
reference surveys rather than relying on a single reference survey.
Because `sp_order = "size"` was used, the reference surveys were ordered
by sample size before estimation. The summary shows that `sp[[2]]`,
corresponding to the second provided probability survey, which is `sp2`
in our example, is the largest reference survey and is therefore treated
as the non-calibrated reference survey in the cumulative precalibration
step. The smaller reference survey, `sp[[1]]`, corresponding to `sp1`,
is calibrated to match selected weighted totals from `sp2`.

The precalibration summary identifies the variables used to align the
two reference surveys before the final pseudo-weight estimation step. In
this example, the calibrated variables include the total survey weight
and the overlapping variables `agecat`, `race`, and `comorbidity`, after
expansion into model-matrix columns such as `agecat2`, `race2`, and
`comorbidity1`. This means that `sp1` is first adjusted so that its
weighted totals for these overlapping variables agree with the
corresponding totals from `sp2`.

The reference samples summary shows how variables are distributed and
selected across the two reference surveys. The section labelled “Shared
variables” lists the variables available from each reference survey
after factor variables have been converted into model-matrix columns.
The section labelled “Variables used for calculation” shows which
columns are actually used in the final multi-reference calibration
equations. Because `sp2` is larger, the overlapping variables are taken
from `sp2`. Therefore, `agecat`, `race`, `BMI`, `diabetes`, and
`comorbidity` are used from `sp2`, while only the variables unique to
`sp1`, namely `education`, `pros_enlarged`, and `psa_level`, are used
from `sp1`. This reflects the main advantage of the multi-reference
interface: it allows different reference surveys to contribute different
parts of the participation model.

The line beginning “Participation model involves the following
variables” lists the full set of auxiliary variables used to estimate
pseudo-weights for the nonprobability sample. The variable names appear
in their model-matrix form. For example, the factor variable `agecat` is
represented by the indicator columns `agecat2`, `agecat3`, and
`agecat4`, with the omitted category serving as the reference category.
Similarly, `race`, `BMI`, `education`, `diabetes`, `comorbidity`, and
`pros_enlarged` are represented by one or more indicator columns. The
variable `psa_level` is included as a numeric auxiliary variable. In
total, the participation model uses demographic, health, education,
prostate-related, and outcome-related information drawn from both
reference surveys.

The solver diagnostics indicate whether the estimating equations were
solved successfully. In this example, the solver is `nleqslv` with the
Newton method. The termination code is 2, and the message states that
the parameter values are within the tolerance `xtol`. This means that
the solver stopped because successive parameter updates became
sufficiently small. The final maximum absolute estimating equations
value is `6.557e-07`. This value is slightly larger than the default
function-value tolerance of `1e-8`, but it is still small in practical
terms. The printed output therefore reports “converged (tolerance
warning)” because convergence was reached through the parameter-step
criterion rather than the function-value criterion. Together with the
small number of iterations, this suggests that the numerical solution is
acceptable for this illustration. In practice, users should inspect both
the termination message and the final maximum absolute estimating
equations value. A successful termination message, combined with a small
`max|EE|`, supports treating the fitted pseudo-weights as numerically
stable. If `max|EE|` is large, or if the solver stops without
convergence, users should consider adjusting the solver controls,
checking sparse factor levels, or revising the participation model.

The participation model coefficients describe the fitted
pseudo-weighting model. These coefficients are used to construct the
pseudo-weights and should not be interpreted as regression coefficients
from an outcome model. In application, they are mainly diagnostic. Users
can inspect them to confirm that the model was fitted and to check
whether any coefficient is unusually large, which may indicate numerical
instability, sparse categories, or strong imbalance between the
nonprobability sample and the reference surveys.

The final line of the summary reports that 1190 observations were
deleted due to missingness in `sp[[1]]`. This refers to (NUMBER OF ROWS
WITH?) missing values in variables used from the first reference survey
after the participation model was specified. Since the default missing
data handling uses complete cases for participation model variables,
these observations are excluded from the pseudo-weight estimation step.
This message is important because heavy missingness in a reference
survey can reduce the amount of information contributed by that survey
and should be considered when interpreting the results.

```{r print-multi-ref-fit}
print(fit_multi)
```

The shorter `print()` output summarizes the fitted object. It reports
that the fitted object has class `"pw_fit"`, uses the multi-reference
calibration method, and contains a participation model with 18
parameters. It also reports the convergence status and the distribution
of the estimated pseudo-weights. In this example, the pseudo-weights
range from 33,303 to 1,099,805, with a median of 100,305 and a mean of
106,180. The sum of pseudo-weights is 255,256,038, which represents the
estimated population size implied by the pseudo-weighted nonprobability
sample. Very large pseudo-weights indicates that some observations in
the nonprobability sample represent population segments that are weakly
represented in the observed sample. This is not necessarily a problem,
but it should be checked because they can increase the variance of
pseudo-weighted estimators.

Taken together, these outputs show that the multi-reference fit uses the
two reference surveys as intended. The larger survey, `sp2`, supplies
the overlapping variables and its survey-specific variables, while `sp1`
supplies the variables not available in `sp2`. The solver diagnostics do
not suggest a serious numerical problem, and the pseudo-weight
distribution gives a useful check on the scale and extremity of the
pseudo-weights. These diagnostics should be checked before using the
fitted object for outcome estimation with `pwmean()`.

## Mean and variance estimation

After the pseudo-weights have been estimated by `est_pw()`, the fitted
object can be passed to `pwmean()` for outcome estimation. In this
example, we first estimate the overall mean of `psa_level` using the
multi-reference pseudo-weights.

```{r multi-psa}
multi_psa <- pwmean(fit_multi, y = "psa_level")

print(multi_psa)
```

The printed output reports the pseudo-weighted undefined mean of serum
PSA level as 1.893 ng/mL, with a standard error of 0.019 and a 95%
confidence interval from 1.855 to 1.931. This estimate is obtained after
adjusting the nonprobability sample with auxiliary information
contributed by both reference surveys.

```{r print-multi_psa}
summary(multi_psa)
```

The summary output also reports the unweighted estimate from the
original nonprobability sample. Before adjustment, the unweighted mean
of psa_level is 1.536 ng/mL, with a 95% confidence interval from 1.468
to 1.605. After pseudo-weighting, the estimated mean increases to 1.893
ng/mL, with a 95% confidence interval from 1.855 to 1.931. The upward
shift suggests that the original nonprobability sample underrepresents
individuals who have auxiliary-variable profiles associated with higher
PSA levels. After the sample is adjusted to better match the
auxiliary-variable information from the reference surveys, the estimated
population mean becomes larger.

Domain-specific estimates can be obtained by supplying a factor variable
or 0/1 numeric variable to `zcol`. For example, the following code
estimates the mean of `psa_level` separately within BMI categories.

```{r multi-domain-psa}
multi_psa_bmi <- pwmean(
  fit_multi,
  y    = "psa_level",
  zcol = "BMI"
)
print(multi_psa_bmi)
```

The domain estimates use the same fitted pseudo-weights from
`fit_multi`. The participation model is not refitted for each BMI
category. Instead, `pwmean()` reuses the pseudo-weights and computes a
separate pseudo-weighted mean within each domain. This is useful when
the same adjustment model is intended to support several outcome or
subgroup analyses.

The same function can also be used when the outcome is categorical. In
this case, `pwmean()` reports estimated prevalence for each category.
The following example treats `diabetes` as a categorical outcome.

```{r multi-categorical-outcome}
multi_diabetes <- pwmean(fit_multi, y = "diabetes")
summary(multi_diabetes)
```

For the diabetes outcome, the unweighted estimate from the
nonprobability sample gives a diabetes prevalence of 17.6%. After
pseudo-weighting, the estimated prevalence decreases to 15.7%, with a
95% confidence interval from 15.2% to 16.2%. Equivalently, the estimated
prevalence of not having diabetes increases from 82.4% before adjustment
to 84.3% after adjustment. This result indicates that the nonprobability
sample contains a higher proportion of individuals with diabetes than
would be expected after adjustment to the auxiliary information
represented by the reference surveys.

## Optional: Sensitivity to cumulative precalibration

Cumulative precalibration can be turned off by setting
`precali = FALSE`. This provides a sensitivity analysis for checking
whether the alignment of overlapping variables across reference surveys
noticeably affects the final estimate. If the pseudo-weighted estimates
with and without cumulative precalibration differ substantially, the
analyst should examine whether the reference surveys target the same
population and whether the overlapping variables are measured and coded
consistently.

```{r multi-no-precali, eval=FALSE}
fit_multi_no_precali <- est_pw(
  data = list(sc, ref1_design, ref2_design),
  p_formula = list(
    ~ agecat + race + education + pros_enlarged + comorbidity,
    ~ agecat + race + BMI + diabetes + comorbidity
  ),
  method   = "multi",
  sp_order = "size",
  precali  = FALSE
)

multi_psa_no_precali <- pwmean(fit_multi_no_precali, y = "psa_level")
print(multi_psa_no_precali)
```

## Optional: One-reference calibration as a special case

A one-reference analysis can be viewed as a special case of the
multi-reference analysis. The same basic components are still present: a
nonprobability sample, a probability reference survey, a participation
model, and a fitted pseudo-weight object. The difference is that all
auxiliary information is from a single reference survey. As a result,
there is no ordering of reference surveys, no cumulative precalibration,
and no variables selection across different reference surveys. **DO WE
NEED TO EXPLAIN THE DIFFERENT METHODS AVAILABLE FOR ONE_REFERENCE CASE
HERE?** **please check the following paragraph**

When one probability reference survey is available, `est_pw()` supports
several pseudo-weighting methods, including raking-ratio calibration
(`method = "calibration"`), adjusted logistic propensity
(`method = "ALP"`), and Chen--Li--Wu (`method = "CLW"`). These methods
share the same workflow: users provide the nonprobability sample and the
reference survey design object, estimate pseudo-weights using
`est_pw()`, and obtain outcome estimates using `pwmean()`.

```{r one-ref-fit}
fit_one <- est_pw(
  data      = list(sc, ref1_design),
  p_formula = ~ agecat + race + education + pros_enlarged +
    comorbidity + BMI + diabetes + psa_level,
  method    = "calibration"
)

summary(fit_one)
```

The summary output reflects this simpler structure. The method is
reported as one-reference calibration, and the participation model
contains variables available in both `sc` and `sp1`. The model includes
demographic variables such as `agecat` and `race`, health variables such
as `comorbidity`, `BMI`, and `diabetes`, and variables such as
`education` and `pros_enlarged` that are available from `sp1`. In
contrast to the multi-reference fit, there is no summary of the
reference surveys showing which variables are taken from each survey.

The solver diagnostics indicate satisfactory convergence. The
termination code is 2, and the message states that the function
criterion is near zero. The final maximum absolute estimating equation
value is 2.235e-08, which is very small. Compared with the
multi-reference fit, the one-reference calibration fit converges in
fewer iterations. This is expected because the one-reference estimating
problem is simpler.

After the pseudo-weights have been estimated, the fitted object can be
passed to `pwmean()` in the same way as the multi-reference fitted
object.

```{r pwmean-one_fit}
pwmean(fit_one,y= "psa_level")
```

## Evaluation against the unweighted estimator

The results in Table \@ref(tab:estimator-comparison) show that
pseudo-weighting changes the PSA estimates in the expected direction. To
quantify this improvement, this subsection compares the multi-reference
pseudo-weighted estimator with the unweighted estimator from the
nonprobability sample. The true finite population values, computed from
the underlying finite population, are used as reference targets for the
evaluation.

**DO WE NEED TO SHOW THREE DECIMAL PLACE? IS TWO ENOUGH?** (original
outcome is needed from Dr. Landsman)

```{r estimator-comparison, echo=FALSE}
true_psa_fp <- c(
  "overall" = 1.976,
  "Normal" = 2.52,
  "Overweight" = 1.83,
  "Obese" = 1.70,
  "Morbidly Obese" = 1.90
)

overall_estimates <- multi_psa$estimates
bmi_estimates <- multi_psa_bmi$estimates
bmi_estimates$domain <- sub("^BMI = ", "", bmi_estimates$domain)

comparison_estimates <- rbind(overall_estimates, bmi_estimates)
comparison_estimates$domain[comparison_estimates$domain == "Overall"] <- "overall"
comparison_estimates <- comparison_estimates[
  match(names(true_psa_fp), comparison_estimates$domain),
]

comparison_table <- data.frame(
  Domain = c("Overall", "Normal", "Overweight", "Obese", "Morbidly Obese"),
  `True value` = true_psa_fp,
  `Unweighted estimate (95% CI)` = sprintf(
    "%.3f (%.3f-%.3f)",
    comparison_estimates$unweighted_mean,
    comparison_estimates$unweighted_lower,
    comparison_estimates$unweighted_upper
  ),
  `Unweighted bias` = sprintf(
    "%.1f%%",
    100 * (comparison_estimates$unweighted_mean - true_psa_fp) / true_psa_fp
  ),
  `Pseudo-weighted estimate (95% CI)` = sprintf(
    "%.3f (%.3f-%.3f)",
    comparison_estimates$adjusted_mean,
    comparison_estimates$adjusted_lower,
    comparison_estimates$adjusted_upper
  ),
  `Pseudo-weighted bias` = sprintf(
    "%.1f%%",
    100 * (comparison_estimates$adjusted_mean - true_psa_fp) / true_psa_fp
  ),
  `Error reduction` = sprintf(
    "%.3f%%",
    100 * (
      abs((comparison_estimates$unweighted_mean - true_psa_fp) / true_psa_fp) -
        abs((comparison_estimates$adjusted_mean - true_psa_fp) / true_psa_fp)
    ) / abs((comparison_estimates$unweighted_mean - true_psa_fp) / true_psa_fp)
  ),
  check.names = FALSE
)

numeric_cols <- vapply(comparison_table, is.numeric, logical(1))
comparison_table[numeric_cols] <- lapply(
  comparison_table[numeric_cols],
  round,
  digits = 3
)

if (knitr::is_latex_output()) {
  comparison_table[[4]] <- gsub("%", "\\%", comparison_table[[4]], fixed = TRUE)
  comparison_table[[6]] <- gsub("%", "\\%", comparison_table[[6]], fixed = TRUE)
  comparison_table[[7]] <- gsub("%", "\\%", comparison_table[[7]], fixed = TRUE)
  names(comparison_table) <- kableExtra::linebreak(
    c(
      "Domain",
      "True\nvalue",
      "Un-\nweighted\nestimate\n(95\\% CI)",
      "Un-\nweighted\nbias",
      "Pseudo-\nweighted\nestimate\n(95\\% CI)",
      "Pseudo-\nweighted\nbias",
      "Error\nreduction"
    ),
    align = "c"
  )
}

comparison_table_output <- knitr::kable(
  comparison_table,
  format = if (knitr::is_latex_output()) "latex" else "html",
  caption = paste(
    "Comparison of unweighted and multi-reference pseudo-weighted estimates"
  ),
  align = c("l", rep("c", 6)),
  row.names = FALSE,
  booktabs = TRUE,
  escape = !knitr::is_latex_output()
)

if (knitr::is_latex_output()) {
  table4_widths <- c("1.6cm", "1.1cm", "2.2cm", "1.3cm", "2.2cm", "1.3cm", "1.3cm")
  for (column_index in seq_along(table4_widths)) {
    comparison_table_output <- kableExtra::column_spec(
      comparison_table_output,
      column = column_index,
      width = table4_widths[[column_index]]
    )
  }
  kableExtra::kable_styling(
    comparison_table_output,
    latex_options = "hold_position",
    font_size = 9
  )
} else {
  kableExtra::kable_styling(
    comparison_table_output,
    bootstrap_options = c("striped", "condensed", "responsive"),
    full_width = TRUE
  )
}
```

Relative bias is computed as the estimator minus the true finite
population value, divided by the true finite population value, and is
reported as a percentage. The table shows that the pseudo-weighted
estimator is closer to the true finite population value than the
unweighted estimator for the overall mean and for every BMI domain. For
the overall PSA mean, the unweighted mean is 1.536, compared with the
true finite population value of 1.976. This gives a relative bias of
-22.3%. After adjustment, the pseudo-weighted mean is 1.893, reducing
the relative bias to -4.2%. This corresponds to an 81.0% reduction in
relative bias.

The same pattern appears across BMI categories. The largest improvement
is observed for the Normal BMI group. In this domain, the relative bias
decreases from -32.8% to -1.5%, corresponding to a 95.4% reduction in
absolute relative bias. The multi-reference estimator also reduces the
magnitude of the relative bias for the Overweight, Obese, and Morbidly
Obese groups, although the size of the improvement differs by domain.

These results show that the gain from pseudo-weighting is not limited to
the overall mean. The adjustment also improves subgroup estimates, even
when the magnitude and direction of the unweighted relative bias differ
across BMI categories. This is relevant for applied analyses because
users often need adjusted estimates not only for the full study
population but also for meaningful subgroups.

Overall, this comparison supports the value of the multi-reference
method in this example. By combining auxiliary information from both
probability reference surveys, the method produces pseudo-weighted
estimates that are closer to the true finite population values than the
unweighted estimates from the nonprobability sample. This provides
empirical evidence that multi-reference calibration method can reduce
bias when variables related to participation are distributed across
multiple probability reference surveys.

# Discussion

This paper introduced `nonprobsampling`, an R package for finite
population inference from a nonprobability sample using auxiliary
information from multiple probability reference surveys. The package
implements pseudo-weighing methods based on the generalized estimating
equations framework of Landsman et al. (2026). The main contribution of
`nonprobsampling` is its support for multi-reference pseudo-weight
estimation with cumulative precalibration. Existing R packages for
nonprobability sample inference mainly support settings in which the
auxiliary variables needed for the participation model come from a
single reference survey. `nonprobsampling` addresses this limitation by
allowing users to combine auxiliary information from multiple reference
surveys within one estimation workflow.

The simulation study showed satisfactory performance of the
multi-reference calibration estimator for category-specific prevalence
estimation across population domains. Across all domains and outcome
categories, relative bias was small, standard error ratios were close to
1, and empirical coverage of 95% confidence intervals was close to the
target level. The estimates obtained with and without cumulative
precalibration were similar in this simulation setting, reflecting the
limited overlap of auxiliary variables between the reference surveys.

The illustration using the package datasets demonstrated the practical
utility of the multi-reference workflow. The pseudo-weighted estimates
were closer to the finite population true values than the unweighted
estimates from the nonprobability sample. For the overall PSA mean, the
relative bias was reduced from -22.3% to -4.2% after pseudo-weighting.
Similar improvements were observed across BMI domains, showing that
multi-reference calibration method can reduce bias not only for overall
estimates but also for subgroup estimates when relevant auxiliary
information is distributed across multiple reference surveys.

Several limitations should be considered. First, the package assumes
that the nonprobability sample and reference surveys are intended to
represent the same finite population. If the reference surveys target a
different population or use incompatible inclusion criteria,
pseudo-weighted estimates may not be meaningful. Second, variables used
in the participation model must be harmonized before estimation. The
package can check names and some data structure issues, but it cannot
determine whether variables are substantively equivalent. Third, the
current user-facing outcome function focuses on means and prevalence.
Regression analyses or hypothesis testing are outside the current main
workflow. Fourth, the package does not impute missing values. It records
and handles missing values according to the selected `na.action`, but
users remain responsible for deciding whether complete-case analysis is
appropriate. Finally, the quality of the resulting pseudo-weighted
estimator depends on the auxiliary variables available in the
nonprobability sample and reference surveys. If important determinants
of participation are absent from all reference surveys, pseudo-weighting
cannot completely remove participation bias. This limitation is not
specific to the software implementation; it is inherent to the use of
reference-survey-based adjustment for nonprobability samples.

From a computational perspective, the current implementation is
primarily intended for typical applied survey analysis settings. Large
samples, high-dimensional participation models, sparse factor levels,
and complex interaction terms can increase memory use and make the
estimating equations more difficult to solve. Users should therefore
inspect solver diagnostics, combine sparse categories, and avoid
unnecessarily complex formulas.

Future extensions could broaden the package in several directions. One
natural extension is to support additional outcome-analysis tools beyond
mean (prevalence) estimation, such as regression models or hypothesis
tests based on fitted pseudo-weights. Another useful direction is to
expand diagnostic tools for evaluating overlap, pseudo-weight
distributions, covariate balance, and sensitivity to the choice of
reference survey ordering. Further work could also consider additional
tools for handling missing data and computational improvements for
larger or more complex analyses. These extensions would build on the
existing two-stage design while preserving the separation between
pseudo-weight estimation and outcome analysis.

Overall, `nonprobsampling` provides a focused software implementation
for a common applied problem: estimating finite population means
(prevalence) from a nonprobability sample when relevant auxiliary
information is distributed across multiple probability reference
surveys. By integrating with the `survey` package, `nonprobsampling`
offers a practical workflow for researchers who need bias-adjusted
estimates from nonprobability samples. It is thus most appropriately
regarded as a complement to current R tools for inference from
nonprobability samples, extending existing workflows by enabling
pseudo-weight estimation using multiple reference surveys and supporting
domain specific estimation.
